---
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
---

<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

#### Load libraries ####
library(usethis)
library(highcharter)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(flextable)

library(htmlwidgets)


#### Add database connection ####
source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("eci_lb_ripa")

## Source the Chart Template
source("chart_functions.R")

```

#### Officer-initiated Stops by Council District and Race
```{r map1, echo=FALSE}
#### Point density map - officer-initiated stops by cd # ####
council_district <- st_read(con, query = "SELECT * FROM data.colb_council_districts_place_2022")
report_stoprates_cd_race_person <- st_read(con, query = "SELECT * FROM data.report_stoprates_cd_race_person")

## make council number a character file
council_district$council_num <- as.character(council_district$council_num) # change council number to character to merge

population_race_cd <-  dbGetQuery(con, "SELECT * FROM data.population_race_cd")

## merge with council district shape-file
council_district_count <-  council_district %>% 
  left_join(
    report_stoprates_cd_race_person
  ) 


# get demographics by council district
df_demographics <- population_race_cd %>% select(cd, race, count, rate) %>%  pivot_wider(
  names_from = race,
  names_glue = "{race}_{.value}",
  values_from = c(count:rate)
)

df_stop <- report_stoprates_cd_race_person %>% select(council_num, race, pop_count, stop_count, stop_rate_per1k) %>%  
  mutate(stop_rate_per1k=round(stop_rate_per1k,1))%>%
  pivot_wider(
 names_from = race,
 names_glue = "{race}_{.value}",
 values_from = c(pop_count:stop_rate_per1k)
)


## merge df_demographics with council_district-- get council_district_demographics
council_district_demographics <- council_district %>% 
  left_join(
    df_demographics, by = c("council_num" = "cd")
  ) 

## merge with stop rate
council_district_demographics <- council_district_demographics %>% left_join(df_stop)


# convert to scattered dots for plotting
lb_ripa_dots = as_dot_density(
  council_district_count,
  value = "stop_count",
  values_per_dot = 10, ## each dot represents 10
  group = "race"
)

## filter for race groups
nh_black_dots <- lb_ripa_dots %>% filter(race == "nh_black")
nh_white_dots <- lb_ripa_dots %>% filter(race == "nh_white")
aian_dots <- lb_ripa_dots %>%  filter(race == "aian")
latinx_dots <- lb_ripa_dots %>% filter(race == "latinx")
mesa_dots <- lb_ripa_dots %>% filter(race == "mesa")
nh_asian_dots <- lb_ripa_dots %>% filter(race == "nh_asian")
nh_twoormor_dots <- lb_ripa_dots %>% filter(race == "nh_twoormor")
nhpi_dots <- lb_ripa_dots %>% filter(race == "nhpi")


## function to change the legend shape and color
colors_function <- function(colors, sizes, borders, shapes) {
  shapes <- gsub("circle", "100%", shapes)
  shapes <- gsub("square", "100%", shapes)
  paste0(colors, "; width:", sizes, "px; height:", sizes, "px; border:1px solid ", borders, "; border-radius:", shapes)
}


### Pop-ups ### ---------------
#create popup
demographic_popup <- paste0("<div class='pointdensity'> <div class='leaflet-data' data-cd='",council_district_demographics$council_num,
               "' data-nh-black-stop-rate='",round(council_district_demographics$nh_black_stop_rate,0),
               "' data-nh-black-rate='",scales::percent(council_district_demographics$nh_black_rate, accuracy = 0.1),
               "' data-latinx-stop-rate='",round(council_district_demographics$latinx_stop_rate,0),
               "' data-latinx-rate='",scales::percent(council_district_demographics$latinx_rate, accuracy = 0.1),
               "' data-nh-asian-stop-rate='",round(council_district_demographics$nh_asian_stop_rate,0),
               "' data-nh-asian-rate='",scales::percent(council_district_demographics$nh_asian_rate, accuracy = 0.1),
               "' data-nh-white-stop-rate='",round(council_district_demographics$nh_white_stop_rate,0),
               "' data-nh-white-rate='",scales::percent(council_district_demographics$nh_white_rate, accuracy = 0.1),
               "' data-mesa-stop-rate='",round(council_district_demographics$mesa_stop_rate,0),
               "' data-mesa-rate='",scales::percent(council_district_demographics$mesa_rate, accuracy = 0.1),
               "' data-nhpi-stop-rate='",round(council_district_demographics$nhpi_stop_rate,0),
               "' data-nhpi-rate='",scales::percent(council_district_demographics$nhpi_rate, accuracy = 0.1),
               "' data-aian-stop-rate='",round(council_district_demographics$aian_stop_rate,0),
               "' data-aian-rate='",scales::percent(council_district_demographics$aian_rate, accuracy = 0.1),
               "' data-nh-twoormor-stop-rate='",round(council_district_demographics$nh_twoormor_stop_rate,0),
               "' data-nh-twoormor-rate='",scales::percent(council_district_demographics$nh_twoormor_rate, accuracy = 0.1),
               "' ></div></div>")



# colors according to our brand guide: https://catalystcalifornia.sharepoint.com/sites/Portal/AP%20Projects/Forms/AllItems.aspx?FolderCTID=0x012000EF0283621C7FE048A8E68512F119353B&id=%2Fsites%2FPortal%2FAP%20Projects%2FCommunications%2FCatalyst%20California%2FBranding%20Materials%20%2B%20Templates%2FVisual%20Brand%20Guide%2FCatalyst%20California%20Brand%20Guide%5FSeptember%202022%2Epdf&viewid=1aaeb4a9%2D2232%2D41c7%2Da71c%2D727af9e1f5d7&parent=%2Fsites%2FPortal%2FAP%20Projects%2FCommunications%2FCatalyst%20California%2FBranding%20Materials%20%2B%20Templates%2FVisual%20Brand%20Guide

# Base Color 00
lime_green <- c('#0AB013')
persian_indigo <- c('#211447')
halloween_orange <- c('#DF4007')
bitter_lemon <- c('#BCD313')
blue <- c('#0860BC')
#dark_tangerine <- c('#EB8D00')
american_yellow <- c('#EABB00')
halaya_ube <- c('#601D42')
dark_silver <- c('#6F6F6F') 



### change size of legend

 leaflet(width = "100%", height = "600px",
        options = leafletOptions(zoomControl = FALSE, minZoom=11.75, maxZoom=11.75)) %>%
   
   # create sidebar and add html for content
addControl(html="<div class='leaflet-popup-scrolled point-density-sidebar'><div class='popup-instruction'><b>Please click on a council district to see stop data below.</b></div><div class ='cd-header'>In Council District <span class='cd' style='font-weight: 600; font-size: 12px; color: #000000;'></span>, LBPD stopped:</div><ul class='popup-list'><li class='popup-list-item'><span class='nh-black-stop-rate nh-black-color'></span> <span class='nh-black-color'>Black people per 1K Black people.</span> They are <span class='nh-black-rate'></span> of CD<span class='cd'></span>'s population.</li><li class='popup-list-item'><span style='color: #000000; font-weight: 400;'><span class='latinx-stop-rate latinx-color'></span> <span class='latinx-color'>Latinx people per 1K Latinx people.</span> They are <span class='latinx-rate'></span> of CD<span class='cd'></span>'s population.</li><li class='popup-list-item'><span style='color: #000000; font-weight: 400;'><span class='nh-asian-stop-rate nh-asian-color'></span> <span class='nh-asian-color'>Asian people per 1K Asian people.</span> They are <span class='nh-asian-rate'></span> of CD<span class='cd'></span>'s population.</li><li class='popup-list-item'><span style='color: #000000; font-weight: 400;'><span class='nh-white-stop-rate nh-white-color'></span> <span class='nh-white-color'>White people per 1K White people.</span> They are <span class='nh-white-rate'></span> of CD<span class='cd'></span>'s population.</li><li class='popup-list-item'><span style='color: #000000; font-weight: 400;'><span class='mesa-stop-rate mesa-color'></span> <span class='mesa-color'>SSWANA people per 1K SSWANA people.</span> They are <span class='mesa-rate'></span> of CD<span class='cd'></span>'s population.</li><li class='popup-list-item'><span style='color: #000000; font-weight: 400;'><span class='nhpi-stop-rate nhpi-color'></span> <span class='nhpi-color'>NHPI people per 1K NHPI people.</span> They are <span class='nhpi-rate'></span> of CD<span class='cd'></span>'s population.</li><li class='popup-list-item'><span style='color: #000000; font-weight: 400;'><span class='aian-stop-rate aian-color'></span> <span class='aian-color'>AIAN people per 1K AIAN people.</span> They are <span class='aian-rate'></span> of CD<span class='cd'></span>'s population.</li><li class='popup-list-item'><span style='color: #000000; font-weight: 400;'><span class='nh-twoormor-stop-rate nh-twoormor-color'></span> <span class='nh-twoormor-color'>people of Two Or More Races per 1K people of Two or More Races.</span> They are <span class='nh-twoormor-rate'></span> of CD<span class='cd'></span>'s population.</li></ul></div>", position = "topleft") %>%

    ### add latinx dots
  
 addCircleMarkers(data = latinx_dots, group = "Latinx", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor= lime_green,  stroke = TRUE) %>% 
   
  ## add NH Black Stops
  
addCircleMarkers(data = nh_black_dots, group = "Black", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor=blue, stroke = TRUE) %>% 
   
   ### add nh asian dots
  
 addCircleMarkers(data = nh_asian_dots, group = "Asian", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor=  halloween_orange, stroke = TRUE) %>% 
   
  
  ## add NH White Stops
  
 addCircleMarkers(data = nh_white_dots, group = "White", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor=  bitter_lemon,  stroke = TRUE) %>% 
   
    ### add MESA dots
  
 addCircleMarkers(data = mesa_dots, group = "MESA", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor= american_yellow, stroke = TRUE) %>% 
  
  
   ### add nhpi dots
  
 addCircleMarkers(data = nhpi_dots, group = "NHPI", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor= halaya_ube, stroke = TRUE) %>% 
   
  ### add AIAN dots 
  
 addCircleMarkers(data = aian_dots,  group = "AIAN", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor= dark_silver, stroke = TRUE) %>% 
  
  ### add nh two or more dots
  
 addCircleMarkers(data = nh_twoormor_dots, group = "Two or More Races", label = NULL, weight = 2, fillOpacity = 1, radius = 2, color="white",  fillColor= persian_indigo, stroke = TRUE) %>% 
  

  
   # add Long beach district layer with stop rates
  addPolygons(data = council_district_demographics, label=~htmlEscape(paste0("Council District ", council_num)), color = "black", fillColor = "white", opacity = 1,  popup = ~paste0(demographic_popup), 
            weight = 2, smoothFactor = 0.5,  highlight = highlightOptions(color = "white", weight = 3, bringToFront = TRUE, sendToBack = FALSE)) %>%
  
  addLayersControl(overlayGroups = c("Latinx","Black",  "Asian", "White",  "MESA", "NHPI", "AIAN", "Two or More Races"),
                   options = layersControlOptions(collapsed = FALSE),
                   position = 'topright')  %>%
  
  #base and view
  
  addProviderTiles("CartoDB.Positron")%>%
  setView(-118.20475, 33.81962, zoom=11.75) %>% #base and view 
  
  hideGroup(c("White"))  %>% # hide white group
 
  htmlwidgets::onRender("
        function() {
            $('.leaflet-control-layers-overlays').append('<label style=\"font-style:italic; font-size:10px;\">*Each dot represents<br>10 people stopped by LBPD.</label>');
        }")
 
invisible(dbDisconnect(con))

```

<style type="text/css">

.info {
    border-radius: 0;
}

.cd-header {
   margin-bottom: 8px; 
   font-weight: 600; 
   color: #000000;
}

.nh-black-color {
  color: #0860BC;
  font-weight: 600;
}

.nh-black-rate {
  color: #000000; 
  font-weight: 400;
  
}

.latinx-color {
  color: #0AB013;
  font-weight: 600;
}

.latinx-rate {
  color: #000000; 
  font-weight: 400;
  
}

.nh-asian-color {
  color: #DF4007;
  font-weight: 600;
}

.nh-asian-rate {
  color: #000000; 
  font-weight: 400;
  
}

.nh-white-color {
  color: #BCD313;
  font-weight: 600;
}

.nh-white-rate {
  color: #000000; 
  font-weight: 400;
  
}

.mesa-color {
  color: #EABB00;
  font-weight: 600;
}

.mesa-rate {
  color: #000000; 
  font-weight: 400;
  
}

.nhpi-color {
  color: #601D42;
  font-weight: 600;
}

.nhpi-rate {
  color: #000000; 
  font-weight: 400;
  
}

.aian-color {
  color: #6F6F6F;
  font-weight: 600;
}

.aian-rate {
  color: #000000; 
  font-weight: 400;
  
}

.nh-twoormor-color {
  color: #211447;
  font-weight: 600;
}

.nh-twoormor-rate {
  color: #000000; 
  font-weight: 400;
  
}

.leaflet-popup-scrolled {
  max-height: 580px;
  opacity: 1;
  border-bottom: 0;
  border-top: 0;
}

.info {
  background: white;
  opacity: 1;
}

.leaflet-left {
  width: 33%;
}

.leaflet-left .leaflet-control {
  font-size: 12px;
  margin-left: 0;
  border-left: 1px solid #ddd;
  border-top: 1px solid #ddd;
  border-bottom: 1px solid #ddd;
}

.leaflet-right .leaflet-control {
  margin-right: 0;
}

.leaflet-top .leaflet-control {
  margin-top: 0;
}

.leaflet-control-container {
  color: #000000;
}

.hidden {
  display: none;
}

.rate-bar {
  height: 1px;
  border: 0;
  border-top: 1px solid;
  margin: 1em 0;
  padding: 0;
  width: 100px;
}
.count-bubble {
  position: relative;
  width: 25px;
  height: 25px;
  border-radius: 50%;
  margin-top: -25px;
  margin-left:100px;
}

.nh-black-visual-color {
  border-top-color: #0860BC;
  background-color: #0860BC;
}

.latinx-visual-color {
  border-top-color: #0AB013;
  background-color: #0AB013;
}

.nh-asian-visual-color {
  border-top-color: #DF4007;
  background-color: #DF4007;
}

.nh-white-visual-color {
  border-top-color: #BCD313;
  background-color: #BCD313;
}

.mesa-visual-color {
  border-top-color: #EABB00;
  background-color: #EABB00;
}

.nhpi-visual-color {
  background-color: #601D42;
  border-top-color: #601D42;
}

.aian-visual-color {
  background-color: #6F6F6F;
  border-top-color: #6F6F6F;
}

.nh-twoormor-visual-color {
  background-color: #211447;
  border-top-color: #211447;
}

.popup-list {
  padding-left: 1.3em;
  line-height: 1.2em;
}

.popup-list-item {
  padding-bottom: 1.2em;
  font-weight: 400; 
  font-size: 1em; 
  line-height: 1.2em;
  color: #000000;
}

.popup-instruction {
  font-style: italic;
  font-weight: 400; 
  font-size: 1em; 
  line-height: 1.2em;
  color: #000000;
  border-bottom: 1px solid #000000;
  padding-top: 1.2em;
  padding-bottom: 1.2em;
  margin-bottom: 1.2em;
}


</style>

```{js}
<!-- This function is used to format LB RIPA data in the point density map "popup"/sidebar feature -->

<!-- Captures  -->
var mapsPlaceholder = [];
L.Map.addInitHook(function () {
  mapsPlaceholder.push(this);
  mapsPlaceholder.forEach(map => {
    map.on('popupopen', (e) => {
    
    let popups = document.querySelectorAll('.leaflet-data');
    console.log(popups);

document.querySelectorAll('.pointdensity').forEach(pointdensity => {
pointdensity.parentElement.parentElement.parentElement.classList.add('hidden')
})
  
popups.forEach((popup) => {
district = popup.getAttribute('data-cd')
document.querySelectorAll('.cd').forEach(node => node.innerHTML = district)
document.querySelector('.nh-black-stop-rate').innerHTML = popup.getAttribute('data-nh-black-stop-rate')
document.querySelector('.nh-black-rate').innerHTML = popup.getAttribute('data-nh-black-rate')
document.querySelector('.latinx-stop-rate').innerHTML = popup.getAttribute('data-latinx-stop-rate')
document.querySelector('.latinx-rate').innerHTML = popup.getAttribute('data-latinx-rate')
document.querySelector('.nh-asian-stop-rate').innerHTML = popup.getAttribute('data-nh-asian-stop-rate')
document.querySelector('.nh-asian-rate').innerHTML = popup.getAttribute('data-nh-asian-rate')
document.querySelector('.nh-white-stop-rate').innerHTML = popup.getAttribute('data-nh-white-stop-rate')
document.querySelector('.nh-white-rate').innerHTML = popup.getAttribute('data-nh-white-rate')
document.querySelector('.mesa-stop-rate').innerHTML = popup.getAttribute('data-mesa-stop-rate')
document.querySelector('.mesa-rate').innerHTML = popup.getAttribute('data-mesa-rate')
document.querySelector('.nhpi-stop-rate').innerHTML = popup.getAttribute('data-nhpi-stop-rate')
document.querySelector('.nhpi-rate').innerHTML = popup.getAttribute('data-nhpi-rate')
document.querySelector('.aian-stop-rate').innerHTML = popup.getAttribute('data-aian-stop-rate')
document.querySelector('.aian-rate').innerHTML = popup.getAttribute('data-aian-rate')
document.querySelector('.nh-twoormor-stop-rate').innerHTML = popup.getAttribute('data-nh-twoormor-stop-rate')
document.querySelector('.nh-twoormor-rate').innerHTML = popup.getAttribute('data-nh-twoormor-rate')

// Rescaling function: (((oldValue - oldMin) * (newMax - newMin)) / (oldMax - oldMin)) + newMin
//let rateScaled = (((stopRate - 0) * (275 - 0)) / (512.1 - 0)) + 5
//let countScaled = (((popCount - 0) * (50 - 10)) / (1 - 0)) + 10
})})})})

```

<div align="left" style="line-height: 1em;"><span style="font-size: 10px;">Race/ethnicity: AIAN=American Indian or Alaska Native, NHPI=Native Hawaiian or Pacific Islander, SSWANA=South Asian, Southwest Asian, or North African; Observations by race may overlap. </span> <br> 
<span style=' font-size: 10px;'>Catalyst California's calculations based on City of Long Beach's Police Stop Data (2019), catalystcalifornia.org, 2023. Analysis for all officer-initiated stops.</span> </div>

